Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-12 2017-Xudong-SMT/FM-2018-extended.tex
Merge documents.
Add the section 6.4 from "checker".
Add the section 6.5 from "elim_var"(never commited to GitHub).

%\usepackage{amsfonts}
%\usepackage{fancyhdr}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{amssymb}
%\usepackage{fancyhdr}
%\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{mathtools,mathpartir}
\usepackage{proof}
%\usepackage{setspace}
%\usepackage{indentfirst}
\usepackage{amsmath,amssymb,amscd,mathrsfs}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{epsfig,color,subfigure,enumitem}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand{\NOTE}[1]{\textcolor{blue}{\textbf{[NOTE:#1]}}}
\newcommand{\ERIC}[1]{\textcolor{blue}{#1}}
\newcommand{\OTvar}{\texttt}
\newcommand{\OTland}{\;\land\ }
\definecolor{airforceblue}{rgb}{0.26, 0.44, 0.56}
\newcommand{\QIN}[1]{\textcolor{airforceblue}{#1}}
\newcommand{\coloncolon}{{:\hspace{-.2ex}:}}
\makeatletter
\newcommand{\raisemath}[1]{\mathpalette{\raisem@th{#1}}}
\newcommand{\raisem@th}[3]{\raisebox{#1}{$#2#3$}}
\makeatother

@@

% Macros for the SOS rules and proof trees:
%\newcommand\openrule[2]{\redinfer{#1}{#2}}
\newcommand\openrule[2]{\inferrule*[myfraction=\reddottedrule,center]{#1}{#2}}
%\newcommand\openrule[2]{\inferrule*{#1}{#2}}
%\newcommand\ostate[1]{\triangleleft{\;#1\;}\triangleright}
\newcommand\ostate[1]{\triangleleft{#1}\triangleright}
\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\usepackage{algorithm} 
%\usepackage{algorithmicx} 
\usepackage{algpseudocode}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\algnewcommand{\IfThen}[2]{\State \algorithmicif\ #1\ \algorithmicthen\ #2}% \IfThenElse{<if>}{<then>}
\algnewcommand{\IfThenElse}[3]{\State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}% \IfThenElse{<if>}{<then>}{<else>}

@@

\begin{algorithm}
\caption{Open Automaton Generation}
\begin{algorithmic}[1]

\State Initialize sets $U$, $E$ for unexplored/explored global states
$U$, $L$ for result OTs;
\While{!isEmpty(U)}
	\State Chose a global state $S$ in $U$; recursively applying Tr1 or Tr2 on the root node;
	\If{Subnet is a pLTS}
                \State Get the local state of the pLTS $s$ from $S$;
		\For{each transition $tr$ starting from $s$} Apply Tr1
			\State Build the corresponding
                        $OT$; store $OT$ in $L$; 
                 \EndFor
	\EndIf
	\If{Subnet is a pNet}
		\For{each $sv \in SV$} Apply Tr2
			\State Invoke $Combining()$ to generate the
                        combination of sub-transitions, $Matching()$
                        to match the combination with the $sv$; Store the result OTs in $L$; 
		\EndFor
	\EndIf
	\For{each target global state $T$ from $OT \in L$}
		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
	\EndFor
\EndWhile
\State Prune the open transitions in $L$ using the SMT solver;
\State \Return $L$;

@@

Algorithm 1 shows the main structure of the algorithm: 
the computation repeats applying Tr1 and Tr2 to build all possible
combinations of behaviors of pLTS and Holes, through all possible
synchronisation vectors. The resulting open transitions are checked
for satisfiability by the SMT solver.

@@

%\paragraph{Combining and Matching using TR2}
Our implementation does not build explicitly the proof trees for each
open transition, but instead : [Combining] enumerate all possible combinations of
pLTS transition (using TR1) and each synchronization vector choice
(TR2); [Matching] build the corresponding transition, and in particular
its predicate, in a hierarchical manner.
Each time we apply Tr2, we build a predicate as a conjunction of two
parts. The first part, which is the simpler one, composes the
predicate from the 
subnets and the guard from transitions in sub-pLTS or synchronization
vectors. The second part is the conjunction of equations generated
during matching the subnets' behaviors with the synchronization
vectors.

@@

\begin{algorithm}
\caption{Combining}
\begin{algorithmic}[1]

%\Require The list $TR$ of the open transition sets of the subnets. 
%\Ensure The list of all the possible combinations of open transitions $L_C$.
\State Initialize an empty combination list $L_C$;
\State Extract a list of open transitions $L_{ot}$ from $TR$;
\State Add $null$ to $L_{ot}$;
\State Get the combinations $L'_C$ of the $TR$;
\For{each $ot \in L_{ot}$}
	\For{each $C' \in L'_C$}
		\State Add $ot$ into $C'$ to get a new tuple $C$;
		\State Add $C$ into $L_C$;
	\EndFor
\EndFor 
\State \Return $L_C$;

\paragraph{Combining:}
Algorithm 2 shows the combining algorithm for enumerating all possible
working status combinations from subnets. Dealing with a list $TR$
containing  sets of open transitions from different subnets,
the algorithm choses one of sets $L_{ot}$ and does combining on the
remaining part of $TR$ recursively. The case when a subnet is not working is
managed by adding a $null$ into the $L_{ot}$.

@@

\begin{algorithm}
\caption{Matching}
\begin{algorithmic}[1]

%\Require 
%The combination of subnets' open transitions $L_C$;
%The behaviors of the holes $B$;
%The set of synchronization vectors $SV$.
%\Ensure 
%The list of all the possible open transitions generated $L_{ot}$.
\State Initial an empty result list $L_{ot}$;
\For{each $C \in L_C$}
	\For{each $sv \in SV$}
		\State Clone the $sv$, the result is $sv'$;
		\State Generate the fresh result action $ra$;
		\State Combine $C, B, ra$ as a new vector $v$;
		\For{each $(e_1 \in v )\&\& (e_2 \in sv')$}
			\If{$(e_1\ is\ null\ \&\&\ e_2\ is\ not\ null)\ ||\ (e_1\ is\ not\ null\ \&\&\ e_2\ is\ null)$} 
			\State Skip;
			\Else ~ Add $(e_1 = e_2)$ into the $Pred$;
			\EndIf
		\EndFor
		\State Generate the result open transition $ot$ with $Pred$;
		\State Add the $ot$ into $L_{ot}$;
	\EndFor
\EndFor 
\State \Return $L_{ot}$;

\paragraph{Matching:} Algorithm 3 shows the algorithm matches
combination $L_C$, hole behaviors $B$ with synchronization vectors
$SV$ according to the definition of the predicate declared before. We
clone the $sv$ using the fresh function to generate a new $sv'$ with
fresh variables.The mismatch on working status is easy to detect so we
filter them in the algorithm through checking the matching elements
$e_1$ and $e_2$. 
Then we build the resulting open transition. Satisfiability of the
predicate is not checked at this level, but only later, at the
toplevel of the pNet.

\subsection{Fresh variables}

For example, in the running example, the first possible behavior of
the hole, it doesn't have a name, only have the prefix ``hb''. The
hole is the second subnets of the root node so its ``tree index'' is
``12''. The fresh variable name is ``$\OTvar{:hb:12:1}$''.

@@

>>>>

%\usepackage{amsfonts}
%\usepackage{fancyhdr}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{amssymb}
%\usepackage{fancyhdr}
%\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{mathtools,mathpartir}
\usepackage{proof}
%\usepackage{setspace}
%\usepackage{indentfirst}
\usepackage{amsmath,amssymb,amscd,mathrsfs}
% \usepackage{array,booktabs,arydshln,xcolor}
\usepackage{array,booktabs,xcolor}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{epsfig,color,subfigure,enumitem}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand{\NOTE}[1]{\textcolor{blue}{\textbf{[NOTE:#1]}}}
\newcommand{\ERIC}[1]{\textcolor{blue}{#1}}
\newcommand{\OTvar}{\texttt}
\newcommand{\OTland}{\;\land\ }
\definecolor{airforceblue}{rgb}{0.26, 0.44, 0.56}
\newcommand{\QIN}[1]{\textcolor{airforceblue}{#1}}
\newcommand{\coloncolon}{{:\hspace{-.2ex}:}}
\makeatletter
\newcommand{\raisemath}[1]{\mathpalette{\raisem@th{#1}}}
\newcommand{\raisem@th}[3]{\raisebox{#1}{$#2#3$}}
\makeatother

@@

% Macros for the SOS rules and proof trees:
%\newcommand\openrule[2]{\redinfer{#1}{#2}}
\newcommand\openrule[2]{\inferrule*[myfraction=\reddottedrule,center]{#1}{#2}}
%\newcommand\openrule[2]{\inferrule*{#1}{#2}}
%\newcommand\ostate[1]{\triangleleft{\;#1\;}\triangleright}
\newcommand\ostate[1]{\triangleleft{#1}\triangleright}
\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\usepackage{algorithm} 
%\usepackage{algorithmicx} 
\usepackage{algpseudocode}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\algnewcommand{\IfThen}[2]{\State \algorithmicif\ #1\ \algorithmicthen\ #2}% \IfThenElse{<if>}{<then>}
\algnewcommand{\IfThenElse}[3]{\State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}% \IfThenElse{<if>}{<then>}{<else>}
%\DeclareMathOperator{\card}{card}
%\DeclareMathOperator{\Flat}{Flat}
\renewcommand{\P}{\mathcal P}
\usepackage{listings}
\lstset{basicstyle=\small}
%\usepackage{algorithm} 
%\usepackage{algpseudocode}

@@

%\begin{algorithm}
%\caption{Open Automaton Generation}
%\begin{algorithmic}[1]
%
%\State Initialize sets $U$, $E$ for unexplored/explored global states
%$U$, $L$ for result OTs;
%\While{!isEmpty(U)}
%	\State Chose a global state $S$ in $U$; recursively applying Tr1 or Tr2 on the root node;
%	\If{Subnet is a pLTS}
%                \State Get the local state of the pLTS $s$ from $S$;
%		\For{each transition $tr$ starting from $s$} Apply Tr1
%			\State Build the corresponding
%                        $OT$; store $OT$ in $L$; 
%                 \EndFor
%	\EndIf
%	\If{Subnet is a pNet}
%		\For{each $sv \in SV$} Apply Tr2
%			\State Invoke $Combining()$ to generate the
%                        combination of sub-transitions, $Matching()$
%                        to match the combination with the $sv$; Store the result OTs in $L$; 
%		\EndFor
%	\EndIf
%	\For{each target global state $T$ from $OT \in L$}
%		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
%	\EndFor
%\EndWhile
%\State Prune the open transitions in $L$ using the SMT solver;
%\State \Return $L$;
%
%\end{algorithmic}  
%\end{algorithm}

\begin{algorithm}
\caption{Open Automaton Generation}
\begin{algorithmic}[1]
\Require The pNet node P.
\State Initialize sets $U$, $E$ for unexplored/explored global states, $L$ for result OTs;
\While{!isEmpty(U)}
	\State Chose $S$ in $U$;
	\State OT = MakeTransitions(P, S);
        \State Store $OT$ in $L$; 
	\For{each target global state $T$ from $OT \in L$}
		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
	\EndFor
\EndWhile
\State Prune the open transitions in $L$ using the SMT solver;
\State \Return $L$;

@@

\begin{algorithm}
\caption{MakeTransitions()}
\begin{algorithmic}[1]
\Require The pNet node P; The start global state S.
\State Initialize a list $l$, $L$ for sub-transitions/transitions.
	\For{each Subnet in P}
	\\ ~~~ $\setminus \setminus$ Recursively applying Tr1 or Tr2 on the sub-nodes.
                \State Store MakeTransitions(Subnet, S) in $l$;
	\EndFor
	\For{each $sv \in SV$}
		\State $comb = Combining(l)$;
                 \State $ot = Matching(sv, comb, hole, v)$;
                 \State Store $ot$ in $L$; 
	\EndFor
\State \Return $L$;
\end{algorithmic}  
\end{algorithm}

\subsection{The Generation of Open Automata}

\QIN{
We have already have the sketch of computing the open automata \cite{???}.
While in our implementation we does not build explicitly a proof tree for every open transition.
Instead, we repeat applying the Tr1 or Tr2 to generate the sub-transitions.
Tr1 is applied on the leaf (pLTS) simply take the pLTS transitions through the given start state and add the predicate in pNet to generate the open transition.
When applying Tr2 we use two methods, combining and matching, to generate the new predicate of the open transitions in a hierarchical manner
as in this case the composition the subnets brings more constraints for synchronizing behaviors. 
Beside the new generated predicate, the predicates from the subnets are also added as a conjunction with those new predicates.
}

%Algorithm 1 shows the main structure of the algorithm: 
%the computation repeats applying Tr1 and Tr2 to build all possible
%combinations of behaviors of pLTS and Holes, through all possible
%synchronisation vectors. The resulting open transitions are checked
%for satisfiability by the SMT solver.

@@

%\paragraph{Combining and Matching using TR2}
%Our implementation does not build explicitly the proof trees for each
%open transition, but instead : [Combining] enumerate all possible combinations of
%pLTS transition (using TR1) and each synchronization vector choice
%(TR2); [Matching] build the corresponding transition, and in particular
%its predicate, in a hierarchical manner.
%Each time we apply Tr2, we build a predicate as a conjunction of two
%parts. The first part, which is the simpler one, composes the
%predicate from the 
%subnets and the guard from transitions in sub-pLTS or synchronization
%vectors. The second part is the conjunction of equations generated
%during matching the subnets' behaviors with the synchronization
%vectors.

@@

%\begin{algorithm}
%\caption{Combining}
%\begin{algorithmic}[1]
%
%%\Require The list $TR$ of the open transition sets of the subnets. 
%%\Ensure The list of all the possible combinations of open transitions $L_C$.
%\State Initialize an empty combination list $L_C$;
%\State Extract a list of open transitions $L_{ot}$ from $TR$;
%\State Add $null$ to $L_{ot}$;
%\State Get the combinations $L'_C$ of the $TR$;
%\For{each $ot \in L_{ot}$}
%	\For{each $C' \in L'_C$}
%		\State Add $ot$ into $C'$ to get a new tuple $C$;
%		\State Add $C$ into $L_C$;
%	\EndFor
%\EndFor 
%\State \Return $L_C$;
%
%\end{algorithmic}  
%\end{algorithm}

\paragraph{Combining:}
\QIN{
The combining method enumerate all the possible status of the subnets.
The result of the method hence call the combination as it present as all the possible combinations of their open transitions.
Assume that there is a collection of $n$ subnets $L = {s_1, s_2, ..., s_n}$, if we notate $ot$ to be the set of open transitions of the subnet $s$ and $\eta$ means the subnet is not involved. 
Then the combination $\texttt{COMB}$, a set of n-tuples, can be computed as an outer product of sets : 
\[\texttt{COMB} : (\eta\cup ot_1) \times (\eta\cup ot_2)\times ... \times (\eta\cup ot_n) .\]
}
%Algorithm 2 shows the combining algorithm for enumerating all possible
%working status combinations from subnets. Dealing with a list $TR$
%containing  sets of open transitions from different subnets,
%the algorithm choses one of sets $L_{ot}$ and does combining on the
%remaining part of $TR$ recursively. The case when a subnet is not working is
%managed by adding a $null$ into the $L_{ot}$.

@@

%\begin{algorithm}
%\caption{Matching}
%\begin{algorithmic}[1]
%%\Require 
%%The combination of subnets' open transitions $L_C$;
%%The behaviors of the holes $B$;
%%The set of synchronization vectors $SV$.
%%\Ensure 
%%The list of all the possible open transitions generated $L_{ot}$.
%\State Initial an empty result list $L_{ot}$;
%\For{each $C \in L_C$}
%	\For{each $sv \in SV$}
%		\State Clone the $sv$, the result is $sv'$;
%		\State Generate the fresh result action $ra$;
%		\State Combine $C, B, ra$ as a new vector $v$;
%		\For{each $(e_1 \in v )\&\& (e_2 \in sv')$}
%			\If{$(e_1\ is\ null\ \&\&\ e_2\ is\ not\ null)\ ||\ (e_1\ is\ not\ null\ \&\&\ e_2\ is\ null)$} 
%			\State Skip;
%			\Else ~ Add $(e_1 = e_2)$ into the $Pred$;
%			\EndIf
%		\EndFor
%		\State Generate the result open transition $ot$ with $Pred$;
%		\State Add the $ot$ into $L_{ot}$;
%	\EndFor
%\EndFor 
%\State \Return $L_{ot}$;
%
%\end{algorithmic} 
%\end{algorithm}

\paragraph{Matching:} 
\QIN{$Given a synchronization vector ~ SV_k\in SV = (a'_i)^{i\in I} (b'_j)^{j\in J}\rightarrow v' , G_k , k\in K$, and a tuple
$C_n\in \texttt{COMB} = {(v_i)}^{i\in I} , n\in N$ where $K$ and $N$ are the indices of the set elements.
As the behavior of subnets should match with the synchronization vector to compose the subnets.
The matching method tries matching $SV_k$ and $C_n$ to generate the predicates.
According to the definition of predicate, the hole behaviors $Hole$ and result action $v$ are also involved. 
}

\QIN{
$$Matching(SV_k, C_n, Hole, v) = 
\begin{array}{l}
\forall {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.
\\
v_i=a'_i\land \, b_j=b'_j \land v=v' \land G_k
\end{array}$$
which is exactly the new part of predicate for one possible result.
}
%Algorithm 3 shows the algorithm matches
%combination $L_C$, hole behaviors $B$ with synchronization vectors
%$SV$ according to the definition of the predicate declared before. We
%clone the $sv$ using the fresh function to generate a new $sv'$ with
%fresh variables.The mismatch on working status is easy to detect so we
%filter them in the algorithm through checking the matching elements
%$e_1$ and $e_2$. 
%Then we build the resulting open transition. Satisfiability of the
%predicate is not checked at this level, but only later, at the
%toplevel of the pNet.

\paragraph{Filtering:} 
\QIN{There is already able to filter some wrong open transitions during the matching.
We can ensure that the synchronization vector is not matched with the subnets if there is only one side not involved.
Then we can remove the transitions from the result of matching according to the condition $\gamma$ where
$\gamma = \bigcup_{i\in I}(v_i\in\eta\land a'_i\in\eta)\lor(v_i\in\eta\land a'_i\in\eta) .$
}

@@

\subsection{Transition to SMTlib}
In order to submit satisfiability problems to Z3 (for the predicates
in open transitions), we need to generate SMTlib programs, from the
pNet Algebra presentation and predicates.
More precisely, we need to translate to SMTlib:
\begin{itemize}
  \item the presentation of the action algebra
    (sorts and operators) that is defined for a given language (process calculus, or high level programming language),
    \item for a given pNet, the set of local constants (actions or auxiliary data) that are used in the pLTSs,
    \item for each open transition: the declaration of variables, and
      the predicate (including action expressions).
\end{itemize}

During this translation, in order to guarantee that the generated code
will cause no runtime errors during parsing and execution, we need
to ensure that all objects used in the SMTlib code are properly
declared, and that they are correctly typed.

Note that in principle, such an algebra correspond to a given
high-level language (e.g. a process algebra), and that the algebra
presentation will be defined once and for all in the framework of the
pNet semantics of each specific language.

\subsubsection{Algebra presentations}

\begin{definition}
  An Algebra Presentation is a pair $\mathcal{P}=<Sorts,Constrs,Ops>$ where:
  \begin{itemize}
  \item $Sorts$ is a set of Sort names
    \footnote{Later we may want to extend this with Sort constructors,
      like Array or Pair, but this is not needed now}
  \item $Constrs$ is a set of constructor operators, with $Con$ the
    constructor name, and arity: $arity(Con)=n \in \mathbb{N}$,
    and with their signature and their associated selectors,
    of the form: '$Con : (sel_1,sort_1), ... , (sel_n,sort_n) -> sort$'.
    For each argument, the pair $(sel_i,sort_i)$ defines an auxiliary
    operator of name $sel_i$ with signature $sel_i : sort -> sort_i$.
    \item $Ops$ is a set of other (auxiliary) operators, with their
      arity and signature, of the form: $Op : sort_1, ...  sort_n ->
      sort$
      \item $Constrs(sortname), Sels(sortname)$, respectively define the set of
        constructors and selectors of sort $sortname$
  \end{itemize}
  All sorts and operator names must be distinct.\\
  Amongst Constructors, those of arity 0 are called constants, and we
  define $Consts(\mathcal{P}) = \{Con \in Constrs. arity(Con)=0\}$.
\end{definition}

We have a minimal, predefined algebra presentation for all pNets, including
three basic sorts $Bool$, $Action$ and $Int$ and their
operators. Table \ref{Table:predefinedSorts} defines these elements.

In addition, and for convenience, we provide one generic construct for
parameterized actions named FUN, which can accept any number of
arguments of any sort. The result type, though, can only be $Action$,
in order to keep the type-checking simple (see next section).

\begin{table}\caption{\label{Table:predefinedSorts}\leftline{Algebra Presentation: predefined Sorts
  and Operators}}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Constructors & Other Operators
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Bool    			&
                $\texttt{true},\ \texttt{false}$&
                $\land,\ \lor$
                \\\specialrule{0em}{1pt}{1pt} 
		Action 			&  $\texttt{FUN}$ &
                \\\specialrule{0em}{1pt}{1pt}
		Int 				&
                ${0, \{i, -i\}_{i \in \texttt{Nat}}}$  &
                $- \texttt{(unary)},\ +,\ -
                \texttt{(binary)},\ \times,\ \div, \texttt{etc.}$
                \\\specialrule{0em}{1pt}{1pt}
                \textsl{for any sort} & & $=,\ \ne$
		\\\hline
	\end{tabular}
\end{table}

For a given language, or for a given use-case, the designer can
declare more sorts and operators, using our pNet API.
As an example, a (value-passing) CCS action algebra, where we assume a
single auxiliary value domain ``Data'', can be defined as:

\begin{example}
  \begin{itemize}
  \item $Sorts_{CCS} = \{Action, Channel, Data, Int, Bool\}$
  \item $Constrs_{CCS} =\\
    \texttt{Emit}: 2, \{(Chan\_E:Channel),(Value\_E:Data)\}->Action\\
    \texttt{Receive}: 2, \{(Chan\_R:Channel),(Value\_R:Data)\}->Action\\
    \texttt{Tau}: 0, \{\}-> Action\\
    \texttt{... and all predefined operators}$
  \end{itemize}
\end{example}

\begin{lstlisting}
AlgebraSort Action = new AlgebraSortImpl("Action");
AlgebraSort Channel = new AlgebraSortImpl("Channel");
AlgebraSort Data = new AlgebraSortImpl("Data");
Action.addConstructor("Emit", 
	{"Chan_E", "Value_E"}, {Channel, Data});
Action.addConstructor("Receive", 
	{"Chan_E", "Value_E"}, {Channel, Data});
Action.addConstructor("Tau");
\end{lstlisting}

\TODO{[DONE][Xudong]: Give an example of defining $Tau$ and $Emit$ using the (new
  version of) the API}

\subsubsection{Expressions}
Once an algebra presentation is defined, we can construct expressions
of the term algebra, using variables and operators:

\begin{definition}
  Let $\mathcal{V}$ be a set of variables. The Term Algebra
  $\Sigma_{\mathcal{P},\mathcal{V}}$ is the set of:
  \begin{itemize}
  \item variables $x \in \mathcal{V}$
  \item terms (or expressions) $op (arg_1, ... , arg_n)$ with $op$ an operator in
  $\mathcal{P}$ of arity $n$, and each $arg_i$ an expression.
  \end{itemize} 
\end{definition}

\paragraph{Constants and variables in pNets and in open transitions}

In addition to the objects defined in the Algebra Presentation, there are
specific objects that are introduced by the pNet construction, and by
the semantic rules used to build the open transitions and their
predicates. This includes, for a pNet $pnet$
\begin{itemize}
\item $Const(pnet)$: Constants from the pLTSs (controllers) transitions: these are
  new constant constructors, usualy of sort Action, local to an
  instance of a controller. The pNet definition requires that all
  these constants are distinct from each other.
\item $SVars(pnet)$: State variables of the controllers. Here also they are required
  to be distinct from those of other controllers.
\item $IVars(pnet)$: Input variables of the controllers.
\item $FVars(pnet,ot)$: Several kinds of ``fresh'' variables, created by application of
  rule Tr2, during the construction of each open transition $ot$:
  Action variables for the behavior of holes and the 
  resulting actions of transitions, variables created during the cloning
  of synchronisation vectors.
\end{itemize}

\def\EPres{\mathcal{P}}
\def\EEnv{\Gamma}

All variable sets above include their sort.
To define the typing rules for expressions and the translation
functions, for any pNet $pnet$ and open transition $ot$, we define an
extended presentation $\EPres_{pnet}$ and environment $\EEnv_{pnet,ot}$ that includes all of the
objects above:

\begin{definition}
  Given an algebra presentation $\mathcal{P}_{pnet}=<Sorts,Constrs,Ops>$ and a pNet $pnet$, we
  construct:
  \begin{itemize}
  \item An extended presentation $\EPres_{pnet}<Sorts,Constrs \cup Const(pnet),Ops>$.
  \item For a given open transition $ot$, an environment
    $\EEnv_{pnet,ot}=SVars(pnet) \cup IVars(pnet) \cup FVars(pnet,ot)$. 
  \end{itemize} 
\end{definition}

\paragraph{Well-formed and well-typed expressions}

The purpose of this section is to define static semantic notions that
will guarantee that the translation to the SMT language will be
correct, i.e. will not yield errors at runtime. This includes
well-formness (all sorts, operators, variables are defined, and
expressions respect the arity of operators), and typing rules.

\begin{definition}
  Given a presentation $\EPres$ (possibly extended) and an environment $\EEnv$:
  \begin{itemize}
  \item $\EEnv$ is well-formed if all sorts in $\EEnv$ are
    defined in $\EPres$
  \item an expression is well-formed if all its operators are defined
    in $\EPres$, and used with the proper arity, and all its
    variables are defined in $\EEnv$
  \item an expression is well-typed if it can be typed by the typing
    rules in table \ref{table:typing-rules}
  \end{itemize}
\end{definition}

The following judgment, and the typing rules in table
\ref{table:typing-rules} can be used to check both the wellformedness
and well-typing of expressions in a pNet or in an open transition,
given the corresponding $\EPres$ and $\EEnv$.

%\begin{table}\caption{\leftline{Typing judgment for Expressions in Open pNets}}
\vspace{1ex}
\noindent
\begin{tabular}{p{5cm}p{7cm}}
		\hline\specialrule{0em}{3pt}{3pt}
%%		$\EEnv \vdash \diamond$ 					& $\EEnv$ is a well-formed environment 					\\\specialrule{0em}{1pt}{1pt}
%%		$\EEnv \vdash A$ 							& A is a well-formed type in $\EEnv$	 					\\\specialrule{0em}{1pt}{1pt}
		$\EPres,\EEnv \vdash M: A$
                & M is a well-formed term of type A in $\EPres,\EEnv$			\\\specialrule{0em}{1pt}{1pt}
		\specialrule{0em}{3pt}{3pt}\hline
	\end{tabular}
%\end{table}

\begin{table}\caption{\leftline{Type Rules for Open pNets}}
  \label{table:typing-rules}
	\begin{tabular}{p{5cm}p{5cm}p{2.5cm}}
		\hline\specialrule{0em}{3pt}{3pt}
		%% (Env $\varnothing$) 								
		%% & 										
		%% &					\\\specialrule{0em}{1pt}{1pt}
            %% $\dfrac{ }{\varnothing \vdash \diamond}$			
            %% & %$\dfrac{\EEnv \vdash A ~~ x\ }{\varnothing \vdash \diamond}$
            %% &					\\\specialrule{0em}{3pt}{3pt}
		%% (Type Bool) 										
		%% &(Type Action) 						
		%% &(Type Int)			\\\specialrule{0em}{1pt}{1pt}
		%% $\dfrac{\EPres \vdash Bool~~\EEnv \vdash \diamond}{\EPres,\EEnv \vdash Bool}$ 
		%% & $\dfrac{\EPres \vdash Action~~\EEnv \vdash \diamond}{\EPres,\EEnv \vdash Action}$ 
		%% & $\dfrac{\EPres \vdash Int~~\EEnv \vdash \diamond}{\EPres,\EEnv \vdash Int}$        \\\specialrule{0em}{3pt}{3pt}
		(Var x) 										
		& 						
		&			\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{\EEnv \vdash x:A}{\EPres,\EEnv \vdash x:A}$ 
		&  
		&       		\\\specialrule{0em}{5pt}{5pt}
		\multispan{2} (Binary operators, e.g.: $\land, \lor$ for booleans, $+, -, \times, \div, \le, \ge$ for integers, etc.)

		& 			\\\specialrule{0em}{3pt}{3pt}
		$\dfrac{\EPres \vdash BinOp :: ty1, ty1 \rightarrow ty2 ~~\EEnv \vdash x_1:ty1 ~~\EEnv \vdash x_2:ty1}{\EPres,\EEnv \vdash x_1 ~BinOp~ x_2:ty2}$ 
		& 
		&       		\\\specialrule{0em}{3pt}{3pt}
		\multicolumn{2}{l}{(Unary operators, e.g. $\lnot$ for booleans, - for integers)}

		& 			\\\specialrule{0em}{3pt}{3pt}
		$\dfrac{\EPres \vdash UnOp :: ty1 \rightarrow ty2 ~~\EEnv \vdash x:ty1}{\EPres,\EEnv \vdash Unop~x:ty2}$   
		&&       		\\\specialrule{0em}{5pt}{5pt}
		(Polymorphic EQ and NEQ)							
		&					
		&
                                         \\\specialrule{0em}{3pt}{3pt}
		$\dfrac{\EPres \vdash A  ~~\EEnv \vdash x_1:A ~~\EEnv \vdash x_2:A}{\EPres,\EEnv \vdash x_1 = x_2:Bool}$ 
		& 
		$\dfrac{\EPres \vdash A  ~~\EEnv \vdash x_1:A ~~\EEnv \vdash x_2:A}{\EPres,\EEnv \vdash x_1 \ne x_2:Bool}$

		&       		\\\specialrule{0em}{5pt}{5pt}
		(Overloaded FUN)							
		&					
		& 			\\\specialrule{0em}{3pt}{3pt}
		$\dfrac{\EPres \vdash \texttt{FUN} :: A_1,...,A_n \rightarrow Action ~~\EPres \vdash A_1~~...~~\EPres \vdash A_n ~~\EEnv \vdash x_1:A_1~~...~~\EEnv \vdash x_n:A_n}{\EPres,\EEnv \vdash \texttt{FUN}(x_1,...,x_n):Action}$ 
		& 
		&			\\
		\specialrule{0em}{5pt}{5pt}\hline
	\end{tabular}
\end{table}

\begin{remark}
  These rules provide a simple type-checking algorithm: if all
  variables in an expression are known in $\EEnv$, then a bottom-up
  application of the rules will decide whether the expression is
  well-typed, and compute the type of each sub-expression.
\end{remark}

\subsubsection{Map to SMT-LIB language}
The pNet elements, as defined above, can be full translated into
SMT-LIB language, but there are a number of differences in the
structure of the models/langages, so the translation is not trivial.

In this section we shall define separately the translation of the
(extended) algebra presentation for one pNet (so it will be fixed for
the study of one use-case); and the translation of each predicate in
the context of an open-transition of this pNet.

\def\translate{~\lhook\joinrel\longrightarrow}

\vspace{1ex}\noindent
%\begin{table}\caption{\leftline{Mapping}}
	\begin{tabular}{p{3cm}p{9cm}}
		\hline\specialrule{0em}{1pt}{1pt}			
		(Presentation)							
		&								\\\specialrule{0em}{1pt}{1pt}		
		&$Sorts \& Constructors \translate$	declare-datatypes				\\\specialrule{0em}{1pt}{1pt}
		&$Operators \lhook\joinrel\longrightarrow$	declare-function		\\\specialrule{0em}{1pt}{1pt}
		(Predicate)							
		&								\\\specialrule{0em}{1pt}{1pt}
		&$\EEnv ~~\lhook\joinrel\longrightarrow$	declare-const		\\\specialrule{0em}{1pt}{1pt}
		&$Pred ~~\lhook\joinrel\longrightarrow$	 assert		\\\specialrule{0em}{1pt}{1pt}
		&$\texttt{FUN}(x_1,...,x_n)$		\\\specialrule{0em}{1pt}{1pt}
		&$\texttt{pNet expression} ~~\lhook\joinrel\longrightarrow$ SMTLib expression	\\\specialrule{0em}{1pt}{1pt}
		&$...$	\\\specialrule{0em}{1pt}{1pt}
		\specialrule{0em}{1pt}{1pt}\hline
	\end{tabular}
%\end{table}

\smallskip
\ERIC{My plan here is to provide the translation principles as some
  high-level pseudo code, together with precise definition of the
  translation functions of a presentation+environment, and of a
  predicate (including expressions)}

In the following definitions, we use abstract functions corresponding to
SMTlib/Z3 constructs. In practice they can be implemented either as
SMTLib scripting programs, or as calls to the Z3 java API.

\subsubsection{Presentation Translation}
We define here the translation of the algebra presentation, extended
with the constant operators collected from the pNet. It produces the
declaration of sorts (excepted Bool and Int) with their constructors
and selectors as \texttt{declare-datatypes}, and the declaration of
other operators as \texttt{declare-function} constructs.
For sorts, we must distinguish the case of mutually defined sorts
(e.g. edges and vertices in a graph), that must be declared within a
single \texttt{declare-datatypes} construct. For this we define:

\begin{definition}
Define the (strict) order "is-using" between sorts $S1 is-using S2$ iff $S2$
occurs as the sort of one argument in the constructors of $S1$.
\end{definition}

\TODO{[Xudong: this is certainly obsolete?] At the same time, we have
  several internal lists (actually hash 
  maps) \texttt{exprs}, \texttt{funcDecls}, \texttt{sortDatatypes} to
  store the generated Z3 objects with their names for later proofs.}
